---
title: "R Notebook"
output: 
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
    code_folding: hide
    keep_md: true
    number_sections: true
---

```{r}
library(here) 

# Set global knitr chunk options
knitr::opts_chunk$set(
    error     = TRUE,        # Ensures that error messages are being shown in the output
    fig.align = "center",    # Ensures that all figures are aligned to the centre in the output
    message   = TRUE,       # <--- Recommend FALSE for cleaner reports, set to TRUE for development
    warning   = TRUE,       # <--- Recommend FALSE for cleaner reports, set to TRUE for development
    autodep   = TRUE,        # Automatically detects dependencies for caching
    cache     = TRUE,        # <--- IMPORTANT: Enable caching globally here!
    results   = "markup",    # Displays the output as normal formatted text
    echo      = TRUE,        # Ensures that the R code itself is displayed in the document
    dev       = "svg",       # Graphics output will be in Scalable Vector Graphics (SVG) format
    out.width = "90%",       # Sets the width of the output image or plot to 90% of the available width 
    out.extra = "keepaspectratio=true" # Adds extra options to the output, preserving aspect ratio
)
```


```{r}
library(here)
```
# Set Up Working Environment

## Set Seed

```{r}
seed <- "210825"
set.seed(seed)
message("seed has been set to 210825 ")
```

## Load Packages & Functions

```{r}
source(here("Custom R Functions and Scripts", "init_packages.R"))
```

## Parallelisation 

```{r}
# Set up parallel processing for CellChat or other parallelized tasks
library(future)
plan("multisession", workers = 10)
options(future.globals.maxSize = 10240 * 1024^2)  # 10 GB memory cap

```

## Configuration

### Custom Colours

```{r}
# Custom colors
man_cols <- c("#86b0cc",  "#f3e65d", "#d5c1e7", "#eeb84c", "#82c39e", "#525252","#4d9f6b", "#b3939e", "#e76031", "#e9944b")
names(man_cols) <- c("B_cells", "NK_cells", "monocytes", "T_cells", "neutrophils", "megakaryocytes", "pDCs",
                     "plasma_cells", "progenitor_cells", "NKT_cells")
```

## Output Folders

### Primary

```{r}
dir.create(here("R Projects", "2907", "output"), recursive = TRUE, showWarnings = FALSE)
```

### DGE

```{r}
dir.create(here("R Projects", "2907", "output", "de_analysis", "Find Markers"), recursive = TRUE, showWarnings = FALSE)
dir.create(here("R Projects", "2907", "output", "de_analysis", "Presto"), recursive = TRUE, showWarnings = FALSE)
dir.create(here("R Projects", "2907", "output", "de_analysis", "t.test"), recursive = TRUE, showWarnings = FALSE)
```

### Visualisation

```{r}
dir.create(here("R Projects", "2907", "output", "visualisation", "box plots"), recursive = TRUE, showWarnings = FALSE)
dir.create(here("R Projects", "2907", "output", "visualisation", "complex heatmaps"), recursive = TRUE, showWarnings = FALSE)
dir.create(here("R Projects", "2907", "output", "visualisation", "nebulosa"), recursive = TRUE, showWarnings = FALSE)
dir.create(here("R Projects", "2907", "output", "visualisation", "violin plots"), recursive = TRUE, showWarnings = FALSE)
dir.create(here("R Projects", "2907", "output", "visualisation", "volcano"), recursive = TRUE, showWarnings = FALSE)
```

## Import & Pre-processing of Data

### Load RDS file

#### Load Original RDS

Importing PhD Clara Savary's RDS & Initialising the object as a Seurat Object

```{r, eval=FALSE}

seu <- readRDS(here("Data",          "seu_NKT_blood_mouse_PPK_ONC_harmony_integrated_filtered_v2_20231_20236_annot.rds"))
```

#### Subset

Initial Subset to isolate T Cells, Natural Killer T Cells and Natural Killer Cells and label

```{r, eval=FALSE}
seu_NKT <- subset(seu, subset = cell_types %in% c("T_cells", "NK_cells", "NKT_cells"))
```

#### Subset Dataset Again

Second Subset -\> subsetting to remove treatment conditions involving theraputics as this is outside the bounds of the scope of the investigation.

Including:

-   DMG Mice (represented as untreated or UT)

-   Control: No Cancer (Treatment and Disease Naive Group)

-   Surgery Control (SHAM - Injected with Saline - to make sure that the stress of surgery isn't a factor and can be compared between Naive, UT and Sham)

Excluding:

-   ONC201 - Dordaviprone -\> DRD2 antagonist and ClpP agonist

    -   <https://doi.org/10.1158/0008-5472.CAN-23-0186>

    -   <https://doi.org/10.1158/2159-8290.CD-23-0131>

-   Dex - Dexmethasone -\> Corticosteroid

-   Dex & ONC201 Combo

```{r, eval=FALSE}
seu_NKT_focused <- subset(seu_NKT, subset = treatment %in% c("UT", "NAIVE", "SHAM"))
```

For later scripting - load seu_NKT_focused

#### Export

Use either base or BioCgenerics

##### BioCgenerics

```{r,eval=FALSE}
BiocGenerics::saveRDS(seu_NKT_focused, file = here("Data", "seu_NKT_focused.rds"))
```

OR

##### Base

```{r, , eval=FALSE}
base::saveRDS(seu_NKT_focused, file = here("Data", "seu_NKT_focused.rds") )
```

#### Load rds

```{r, cache=FALSE}
seu_NKT_focused <- readRDS(here("Data", "seu_NKT_focused.rds"))
```



### T Cell Subset 

```{r, cache=FALSE}
seu_T <- subset(seu_NKT_focused, subset = T_clusters %in% "T_cells")
```
# QC

Already completed by PhD CS


# Cell Type Annotation

```{r}

```


# Dim Reduction & Clustering

# Differential Expression Analysis 
MAST, Find Markers, Presto and deseq2/edgeR


# Receptor Localistion & Surface vs Total Expression



# Ligand–Receptor Interaction Analysis

# Trajectory Analysis/Pseudotime
```{r, cache=TRUE, eval=FALSE}
cds <- as.cell_data_set(seu_NKT_focused)
cds <- preprocess_cds(cds, num_dim = 50)
cds <- cluster_cells(cds, reduction_method = "UMAP")  # <-- fix here
cds <- reduce_dimension(cds, reduction_method = "UMAP")
cds <- learn_graph(cds)
plot_cells(cds, color_cells_by = "pseudotime")

```


# Vis

## Dittoplot

```{r}
dittoPlot(seu_T, "S1pr1", group.by = "treatment", split.by = "tissue")
```

```{r}
dittoBoxPlot(seu_T, "S1pr1", group.by = "treatment", split.by = "tissue")
```
```{r}
dittoRidgePlot(seu_T, "S1pr1", group.by = "treatment", split.by = "tissue")
```

## Enhanced Volcano Plots

## Nebulosa




