```{r}
knitr::opts_chunk$set(
    error     = TRUE,     # Ensures that error messages are being shown in the output
    fig.align = "center", # Ensures that all figures are aligned to the centre in the output
    message   = TRUE,     # Display messages generated by R
    warning   = TRUE,     # Display warnings generated by R
    autodep   = TRUE,     # Automatically detects dependencies between chunks and reruns dependent chunks when needed
    cache     = FALSE,    # Disables caching, meaning every time the document is knit, all chunks are re-executed
    results   = "markup", # Displays the output as normal formatted text
    echo      = TRUE     # Ensures that the R code itself is displayed in the document
    #dev       = "svg",    # Graphics output will be in Scalable Vector Graphics (SVG) format
    #out.width = "90%",    # Sets the width of the output image or plot to 90% of the available width 
  #out.extra = "keepaspectratio=true" # Adds extra options to the output, in this case, preserving the aspect ratio of the image or plot
)
```

```{r}
# Maximum allowable size for globals in bytes that can be exported to the parallel workers
mib <- 9000*1024^2
options(future.globals.maxSize = mib)  # 1.5 GiB
```

```{r}
library(here)
```

# Set Up Working Environment

## Set Seed

```{r}
set.seed(210825)
```

## Load Packages & Functions

```{r}
source(here("Custom R Functions and Scripts", "init_packages.R"))
```

# Configuration

## Custom Colours

```{r}
# Custom colors
man_cols <- c("#86b0cc",  "#f3e65d", "#d5c1e7", "#eeb84c", "#82c39e", "#525252","#4d9f6b", "#b3939e", "#e76031", "#e9944b")
names(man_cols) <- c("B_cells", "NK_cells", "monocytes", "T_cells", "neutrophils", "megakaryocytes", "pDCs",
                     "plasma_cells", "progenitor_cells", "NKT_cells")
```

## Output Folders

### Primary

```{r}
dir.create(here("R Projects", "2107", "output"))
```

### DGE

```{r}
dir.create(here("R Projects", "2107", "output", "de_analysis"))
```

# Import & Pre-processing of Data

## Load RDS file

### Load Original RDS

Importing PhD Clara Savary's RDS & Initialising the object as a Seurat Object

```{r}

seu <- readRDS(here("Data",          "seu_NKT_blood_mouse_PPK_ONC_harmony_integrated_filtered_v2_20231_20236_annot.rds"))
```

### Subset

Initial Subset to isolate T Cells, Natural Killer T Cells and Natural Killer Cells and label

```{r}
seu_NKT <- subset(seu, subset = cell_types %in% c("T_cells", "NK_cells", "NKT_cells"))
```

### Subset Dataset Again

Second Subset -\> subsetting to remove treatment conditions involving theraputics as this is outside the bounds of the scope of the investigation.

Including:

-   DMG Mice (represented as untreated or UT)

-   Control: No Cancer (Treatment and Disease Naive Group)

-   Surgery Control (SHAM - Injected with Saline - to make sure that the stress of surgery isn't a factor and can be compared between Naive, UT and Sham)

Excluding:

-   ONC201 - Dordaviprone -\> DRD2 antagonist and ClpP agonist

    -   <https://doi.org/10.1158/0008-5472.CAN-23-0186>

    -   <https://doi.org/10.1158/2159-8290.CD-23-0131>

-   Dex - Dexmethasone -\> Corticosteroid

-   Dex & ONC201 Combo

```{r}
seu_NKT_focused <- subset(seu_NKT, subset = treatment %in% c("UT", "NAIVE", "SHAM"))
```

For later scripting - load seu_NKT_focused

#### Export

Use either base or Biogenerics

##### Biogenerics

```{r}
saveRDS(seu_NKT_focused, file = here("Data", "seu_NKT_focused") )
```

##### Base

```{r}
base::saveRDS(seu_NKT_focused, file = here("Data", "seu_NKT_focused") )
```

#### Load RDS

```{r}
seu_NKT_focused <- readRDS(here("Data", "seu_NKT_focused.RDS"))
```

# Differential Gene Expression Analysis

## T Cells

```{r}
seu_T <- subset(seu_NKT_focused, subset = T_clusters %in% "T_cells")
```

### (Seurat::Find Markers:wilcox Limma)

```{r}
# Define the gene list you want to focus on
selected_genes <- c("S1pr1", "S1pr2", "S1pr3", "S1pr4", "S1pr5", 
                    "Ccr7", "Sell", "Cd69", "Klf2", "Cxcr4", "Cd44", "Itgae", "Itgal")

pairwise_comparisons <- list(
  "NAIVE_WB_vs_NAIVE_BM"     = c("Naive_WB", "Naive_BM"),
  "NAIVE_WB_vs_SHAM_WB"      = c("Naive_WB", "Sham_WB"),
  "NAIVE_WB_vs_SHAM_BM"      = c("Naive_WB", "Sham_BM"),
  "NAIVE_WB_vs_UT_WB"        = c("Naive_WB", "UT_WB"),
  "NAIVE_WB_vs_UT_BM"        = c("Naive_WB", "UT_BM"),
  "NAIVE_BM_vs_SHAM_WB"      = c("Naive_BM", "Sham_WB"),
  "NAIVE_BM_vs_SHAM_BM"      = c("Naive_BM", "Sham_BM"),
  "NAIVE_BM_vs_UT_WB"        = c("Naive_BM", "UT_WB"),
  "NAIVE_BM_vs_UT_BM"        = c("Naive_BM", "UT_BM"),
  "SHAM_WB_vs_SHAM_BM"       = c("Sham_WB", "Sham_BM"),
  "SHAM_WB_vs_UT_WB"         = c("Sham_WB", "UT_WB"),
  "SHAM_WB_vs_UT_BM"         = c("Sham_WB", "UT_BM"),
  "SHAM_BM_vs_UT_WB"         = c("Sham_BM", "UT_WB"),
  "SHAM_BM_vs_UT_BM"         = c("Sham_BM", "UT_BM"),
  "UT_WB_vs_UT_BM"           = c("UT_WB", "UT_BM")
)

wb <- createWorkbook()

for (i in names(pairwise_comparisons)) {
  
  deg_T <- FindMarkers(seu_T, group.by = "sample_name", ident.1 = pairwise_comparisons[[i]][1], ident.2 = pairwise_comparisons[[i]][2], test.use = "wilcox_limma")
  
  addWorksheet(wb, sheetName = i)
  
  deg_T <- deg_T %>%
    filter(rownames(.) %in% selected_genes) %>%   # <- Filter only your selected genes
    arrange(-avg_log2FC)
  
  # Add gene names as a column
  deg_T$gene <- rownames(deg_T)
  deg_T <- deg_T %>% select(gene, everything())
  
  writeData(wb, sheet = i, x = deg_T)
}

saveWorkbook(wb, here("R Projects", "2107", "output", "de_analysis", "DE_wilcox-limma_filtered.xlsx"), overwrite = TRUE)

```

### (Presto) Wilcoxon signed-rank test

```{r}

# Define the gene list you want to focus on
selected_genes <- c("S1pr1", "S1pr2", "S1pr3", "S1pr4", "S1pr5", 
                    "Ccr7", "Sell", "Cd69", "Klf2", "Cxcr4", "Cd44", "Itgae", "Itgal")

pairwise_comparisons <- list(
  "NAIVE_WB_vs_NAIVE_BM"     = c("Naive_WB", "Naive_BM"),
  "NAIVE_WB_vs_SHAM_WB"      = c("Naive_WB", "Sham_WB"),
  "NAIVE_WB_vs_SHAM_BM"      = c("Naive_WB", "Sham_BM"),
  "NAIVE_WB_vs_UT_WB"        = c("Naive_WB", "UT_WB"),
  "NAIVE_WB_vs_UT_BM"        = c("Naive_WB", "UT_BM"),
  "NAIVE_BM_vs_SHAM_WB"      = c("Naive_BM", "Sham_WB"),
  "NAIVE_BM_vs_SHAM_BM"      = c("Naive_BM", "Sham_BM"),
  "NAIVE_BM_vs_UT_WB"        = c("Naive_BM", "UT_WB"),
  "NAIVE_BM_vs_UT_BM"        = c("Naive_BM", "UT_BM"),
  "SHAM_WB_vs_SHAM_BM"       = c("Sham_WB", "Sham_BM"),
  "SHAM_WB_vs_UT_WB"         = c("Sham_WB", "UT_WB"),
  "SHAM_WB_vs_UT_BM"         = c("Sham_WB", "UT_BM"),
  "SHAM_BM_vs_UT_WB"         = c("Sham_BM", "UT_WB"),
  "SHAM_BM_vs_UT_BM"         = c("Sham_BM", "UT_BM"),
  "UT_WB_vs_UT_BM"           = c("UT_WB", "UT_BM")
)

wb <- createWorkbook()

for (i in names(pairwise_comparisons)) {
  
  deg_T <- wilcoxauc(seu_T, group_by = "sample_name", groups_use = pairwise_comparisons[[i]])
  
  addWorksheet(wb, sheetName = i)
  
  deg_T <- deg_T %>%
    filter(group == pairwise_comparisons[[i]][[1]]) %>%
    filter(feature %in% selected_genes) %>%   # <- Filter only your selected genes
    arrange(group, -logFC)
  
  writeData(wb, sheet = i, x = deg_T)
}

saveWorkbook(wb, here("R Projects", "2107", "output", "de_analysis", "DE_presto-wilcoxauc_filtered.xlsx"), overwrite = TRUE)

```

### Wilcox Presto Find Markers

```{r}
# Define the gene list you want to focus on
selected_genes <- c("S1pr1", "S1pr2", "S1pr3", "S1pr4", "S1pr5", 
                    "Ccr7", "Sell", "Cd69", "Klf2", "Cxcr4", "Cd44", "Itgae", "Itgal")

pairwise_comparisons <- list(
  "NAIVE_WB_vs_NAIVE_BM"     = c("Naive_WB", "Naive_BM"),
  "NAIVE_WB_vs_SHAM_WB"      = c("Naive_WB", "Sham_WB"),
  "NAIVE_WB_vs_SHAM_BM"      = c("Naive_WB", "Sham_BM"),
  "NAIVE_WB_vs_UT_WB"        = c("Naive_WB", "UT_WB"),
  "NAIVE_WB_vs_UT_BM"        = c("Naive_WB", "UT_BM"),
  "NAIVE_BM_vs_SHAM_WB"      = c("Naive_BM", "Sham_WB"),
  "NAIVE_BM_vs_SHAM_BM"      = c("Naive_BM", "Sham_BM"),
  "NAIVE_BM_vs_UT_WB"        = c("Naive_BM", "UT_WB"),
  "NAIVE_BM_vs_UT_BM"        = c("Naive_BM", "UT_BM"),
  "SHAM_WB_vs_SHAM_BM"       = c("Sham_WB", "Sham_BM"),
  "SHAM_WB_vs_UT_WB"         = c("Sham_WB", "UT_WB"),
  "SHAM_WB_vs_UT_BM"         = c("Sham_WB", "UT_BM"),
  "SHAM_BM_vs_UT_WB"         = c("Sham_BM", "UT_WB"),
  "SHAM_BM_vs_UT_BM"         = c("Sham_BM", "UT_BM"),
  "UT_WB_vs_UT_BM"           = c("UT_WB", "UT_BM")
)

wb <- createWorkbook()

for (i in names(pairwise_comparisons)) {
  
  deg_T <- FindMarkers(seu_T, group.by = "sample_name", ident.1 = pairwise_comparisons[[i]][1], ident.2 = pairwise_comparisons[[i]][2], test.use = "wilcox")
  
  addWorksheet(wb, sheetName = i)
  
  deg_T <- deg_T %>%
    filter(rownames(.) %in% selected_genes) %>%   # <- Filter only your selected genes
    arrange(-avg_log2FC)
  
  # Add gene names as a column
  deg_T$gene <- rownames(deg_T)
  deg_T <- deg_T %>% select(gene, everything())
  
  writeData(wb, sheet = i, x = deg_T)
}

saveWorkbook(wb, here("R Projects", "2107", "output", "de_analysis", "DE_wilcox-presto(seurat)_filtered.xlsx"), overwrite = TRUE)

```

### deseq2

```{r}
# Define the gene list you want to focus on
selected_genes <- c("S1pr1", "S1pr2", "S1pr3", "S1pr4", "S1pr5", 
                    "Ccr7", "Sell", "Cd69", "Klf2", "Cxcr4", "Cd44", "Itgae", "Itgal")

pairwise_comparisons <- list(
  "NAIVE_WB_vs_NAIVE_BM"     = c("Naive_WB", "Naive_BM"),
  "NAIVE_WB_vs_SHAM_WB"      = c("Naive_WB", "Sham_WB"),
  "NAIVE_WB_vs_SHAM_BM"      = c("Naive_WB", "Sham_BM"),
  "NAIVE_WB_vs_UT_WB"        = c("Naive_WB", "UT_WB"),
  "NAIVE_WB_vs_UT_BM"        = c("Naive_WB", "UT_BM"),
  "NAIVE_BM_vs_SHAM_WB"      = c("Naive_BM", "Sham_WB"),
  "NAIVE_BM_vs_SHAM_BM"      = c("Naive_BM", "Sham_BM"),
  "NAIVE_BM_vs_UT_WB"        = c("Naive_BM", "UT_WB"),
  "NAIVE_BM_vs_UT_BM"        = c("Naive_BM", "UT_BM"),
  "SHAM_WB_vs_SHAM_BM"       = c("Sham_WB", "Sham_BM"),
  "SHAM_WB_vs_UT_WB"         = c("Sham_WB", "UT_WB"),
  "SHAM_WB_vs_UT_BM"         = c("Sham_WB", "UT_BM"),
  "SHAM_BM_vs_UT_WB"         = c("Sham_BM", "UT_WB"),
  "SHAM_BM_vs_UT_BM"         = c("Sham_BM", "UT_BM"),
  "UT_WB_vs_UT_BM"           = c("UT_WB", "UT_BM")
)

wb <- createWorkbook()

for (i in names(pairwise_comparisons)) {
  
  deg_T <- FindMarkers(seu_T, group.by = "sample_name", ident.1 = pairwise_comparisons[[i]][1], ident.2 = pairwise_comparisons[[i]][2], test.use = "DESeq2")
  
  addWorksheet(wb, sheetName = i)
  
  deg_T <- deg_T %>%
    filter(rownames(.) %in% selected_genes) %>%   # <- Filter only your selected genes
    arrange(-avg_log2FC)
  
  # Add gene names as a column
  deg_T$gene <- rownames(deg_T)
  deg_T <- deg_T %>% select(gene, everything())
  
  writeData(wb, sheet = i, x = deg_T)
}

saveWorkbook(wb, here("R Projects", "2107", "output", "de_analysis", "DE_deseq2_filtered.xlsx"), overwrite = TRUE)

```

### edgeR - Empirical Methods for DE (Negative Binomial Distribution)

```{r}
selected_genes <- c("S1pr1", "S1pr2", "S1pr3", "S1pr4", "S1pr5", 
                    "Ccr7", "Sell", "Cd69", "Klf2", "Cxcr4", "Cd44", "Itgae", "Itgal")

pairwise_comparisons <- list(
  "NAIVE_WB_vs_NAIVE_BM"     = c("Naive_WB", "Naive_BM"),
  "NAIVE_WB_vs_SHAM_WB"      = c("Naive_WB", "Sham_WB"),
  "NAIVE_WB_vs_SHAM_BM"      = c("Naive_WB", "Sham_BM"),
  "NAIVE_WB_vs_UT_WB"        = c("Naive_WB", "UT_WB"),
  "NAIVE_WB_vs_UT_BM"        = c("Naive_WB", "UT_BM"),
  "NAIVE_BM_vs_SHAM_WB"      = c("Naive_BM", "Sham_WB"),
  "NAIVE_BM_vs_SHAM_BM"      = c("Naive_BM", "Sham_BM"),
  "NAIVE_BM_vs_UT_WB"        = c("Naive_BM", "UT_WB"),
  "NAIVE_BM_vs_UT_BM"        = c("Naive_BM", "UT_BM"),
  "SHAM_WB_vs_SHAM_BM"       = c("Sham_WB", "Sham_BM"),
  "SHAM_WB_vs_UT_WB"         = c("Sham_WB", "UT_WB"),
  "SHAM_WB_vs_UT_BM"         = c("Sham_WB", "UT_BM"),
  "SHAM_BM_vs_UT_WB"         = c("Sham_BM", "UT_WB"),
  "SHAM_BM_vs_UT_BM"         = c("Sham_BM", "UT_BM"),
  "UT_WB_vs_UT_BM"           = c("UT_WB", "UT_BM")
)

wb <- createWorkbook()

for (i in names(pairwise_comparisons)) {
  
  deg_T <- FindMarkers(seu_T, group.by = "sample_name", ident.1 = pairwise_comparisons[[i]][1], ident.2 = pairwise_comparisons[[i]][2], test.use = "negbinom")
  
  addWorksheet(wb, sheetName = i)
  
  deg_T <- deg_T %>%
    filter(rownames(.) %in% selected_genes) %>%   # <- Filter only your selected genes
    arrange(-avg_log2FC)
  
  # Add gene names as a column
  deg_T$gene <- rownames(deg_T)
  deg_T <- deg_T %>% select(gene, everything())
  
  writeData(wb, sheet = i, x = deg_T)
}

saveWorkbook(wb, here("R Projects", "2107", "output", "de_analysis", "DE_edger_filtered.xlsx"), overwrite = TRUE)


```

### MAST

```{r}
# Define the gene list you want to focus on
selected_genes <- c("S1pr1", "S1pr2", "S1pr3", "S1pr4", "S1pr5", 
                    "Ccr7", "Sell", "Cd69", "Klf2", "Cxcr4", "Cd44", "Itgae", "Itgal")

pairwise_comparisons <- list(
  "NAIVE_WB_vs_NAIVE_BM"     = c("Naive_WB", "Naive_BM"),
  "NAIVE_WB_vs_SHAM_WB"      = c("Naive_WB", "Sham_WB"),
  "NAIVE_WB_vs_SHAM_BM"      = c("Naive_WB", "Sham_BM"),
  "NAIVE_WB_vs_UT_WB"        = c("Naive_WB", "UT_WB"),
  "NAIVE_WB_vs_UT_BM"        = c("Naive_WB", "UT_BM"),
  "NAIVE_BM_vs_SHAM_WB"      = c("Naive_BM", "Sham_WB"),
  "NAIVE_BM_vs_SHAM_BM"      = c("Naive_BM", "Sham_BM"),
  "NAIVE_BM_vs_UT_WB"        = c("Naive_BM", "UT_WB"),
  "NAIVE_BM_vs_UT_BM"        = c("Naive_BM", "UT_BM"),
  "SHAM_WB_vs_SHAM_BM"       = c("Sham_WB", "Sham_BM"),
  "SHAM_WB_vs_UT_WB"         = c("Sham_WB", "UT_WB"),
  "SHAM_WB_vs_UT_BM"         = c("Sham_WB", "UT_BM"),
  "SHAM_BM_vs_UT_WB"         = c("Sham_BM", "UT_WB"),
  "SHAM_BM_vs_UT_BM"         = c("Sham_BM", "UT_BM"),
  "UT_WB_vs_UT_BM"           = c("UT_WB", "UT_BM")
)

wb <- createWorkbook()

for (i in names(pairwise_comparisons)) {
  
  deg_T <- FindMarkers(seu_T, group.by = "sample_name", ident.1 = pairwise_comparisons[[i]][1], ident.2 = pairwise_comparisons[[i]][2], test.use = "MAST")
  
  addWorksheet(wb, sheetName = i)
  
  deg_T <- deg_T %>%
    filter(rownames(.) %in% selected_genes) %>%   # <- Filter only your selected genes
    arrange(-avg_log2FC)
  
  # Add gene names as a column
  deg_T$gene <- rownames(deg_T)
  deg_T <- deg_T %>% select(gene, everything())
  
  writeData(wb, sheet = i, x = deg_T)
}

saveWorkbook(wb, here("R Projects", "2107", "output", "de_analysis", "DE_mast_filtered.xlsx"), overwrite = TRUE)


```

### Student's T Test

```{r}
# Define the gene list you want to focus on
selected_genes <- c("S1pr1", "S1pr2", "S1pr3", "S1pr4", "S1pr5", 
                    "Ccr7", "Sell", "Cd69", "Klf2", "Cxcr4", "Cd44", "Itgae", "Itgal")

pairwise_comparisons <- list(
  "NAIVE_WB_vs_NAIVE_BM"     = c("Naive_WB", "Naive_BM"),
  "NAIVE_WB_vs_SHAM_WB"      = c("Naive_WB", "Sham_WB"),
  "NAIVE_WB_vs_SHAM_BM"      = c("Naive_WB", "Sham_BM"),
  "NAIVE_WB_vs_UT_WB"        = c("Naive_WB", "UT_WB"),
  "NAIVE_WB_vs_UT_BM"        = c("Naive_WB", "UT_BM"),
  "NAIVE_BM_vs_SHAM_WB"      = c("Naive_BM", "Sham_WB"),
  "NAIVE_BM_vs_SHAM_BM"      = c("Naive_BM", "Sham_BM"),
  "NAIVE_BM_vs_UT_WB"        = c("Naive_BM", "UT_WB"),
  "NAIVE_BM_vs_UT_BM"        = c("Naive_BM", "UT_BM"),
  "SHAM_WB_vs_SHAM_BM"       = c("Sham_WB", "Sham_BM"),
  "SHAM_WB_vs_UT_WB"         = c("Sham_WB", "UT_WB"),
  "SHAM_WB_vs_UT_BM"         = c("Sham_WB", "UT_BM"),
  "SHAM_BM_vs_UT_WB"         = c("Sham_BM", "UT_WB"),
  "SHAM_BM_vs_UT_BM"         = c("Sham_BM", "UT_BM"),
  "UT_WB_vs_UT_BM"           = c("UT_WB", "UT_BM")
)

# Create an Excel workbook
wb <- createWorkbook()

for (i in names(pairwise_comparisons)) {
  
  deg_T <- FindMarkers(
    seu_T,
    group.by = "sample_name",
    ident.1 = pairwise_comparisons[[i]][1],
    ident.2 = pairwise_comparisons[[i]][2],
    test.use = "t"
  )
  
  addWorksheet(wb, sheetName = i)
  
  # Filter for selected genes and sort by log2FC
  deg_T <- deg_T %>%
    filter(rownames(.) %in% selected_genes) %>%
    arrange(-avg_log2FC)
  
  # Add gene names as a column and reorder
  deg_T$gene <- rownames(deg_T)
  deg_T <- deg_T %>% dplyr::select(gene, dplyr::everything())
  
  # Write to Excel sheet
  writeData(wb, sheet = i, x = deg_T)
}

# Save the workbook
saveWorkbook(
  wb,
  here("R Projects", "2107", "output", "de_analysis", "DE_Ttest_filtered.xlsx"),
  overwrite = TRUE
)
```
