```{r}
knitr::opts_chunk$set(
    error     = TRUE,     # Ensures that error messages are being shown in the output
    fig.align = "center", # Ensures that all figures are aligned to the centre in the output
    message   = TRUE,     # Display messages generated by R
    warning   = TRUE,     # Display warnings generated by R
    autodep   = TRUE,     # Automatically detects dependencies between chunks and reruns dependent chunks when needed
    cache     = FALSE,    # Disables caching, meaning every time the document is knit, all chunks are re-executed
    results   = "markup", # Displays the output as normal formatted text
    echo      = TRUE     # Ensures that the R code itself is displayed in the document
    #dev       = "svg",    # Graphics output will be in Scalable Vector Graphics (SVG) format
    #out.width = "90%",    # Sets the width of the output image or plot to 90% of the available width 
  #out.extra = "keepaspectratio=true" # Adds extra options to the output, in this case, preserving the aspect ratio of the image or plot
)
```

```{r}
# Maximum allowable size for globals in bytes that can be exported to the parallel workers
mib <- 9000*1024^2
options(future.globals.maxSize = mib)  # 1.5 GiB
```

```{r}
library(here)
```

# Set Up Working Environment

## Load Packages & Functions

```{r}
source(here("Custom R Functions and Scripts", "init_packages.R"))
```

# Configuration

Custom Colours

```{r}
# Custom colors
man_cols <- c("#86b0cc",  "#f3e65d", "#d5c1e7", "#eeb84c", "#82c39e", "#525252","#4d9f6b", "#b3939e", "#e76031", "#e9944b")
names(man_cols) <- c("B_cells", "NK_cells", "monocytes", "T_cells", "neutrophils", "megakaryocytes", "pDCs",
                     "plasma_cells", "progenitor_cells", "NKT_cells")
```

# Import & Pre-processing of Data

## Load RDS file

### Load Original RDS

Importing PhD Clara Savary's RDS & Initialising the object as a Seurat Object

```{r}

seu <- readRDS(here("Data",          "seu_NKT_blood_mouse_PPK_ONC_harmony_integrated_filtered_v2_20231_20236_annot.rds"))
```

### Subset

Initial Subset to isolate T Cells, Natural Killer T Cells and Natural Killer Cells and label

```{r}
seu_NKT <- subset(seu, subset = cell_types %in% c("T_cells", "NK_cells", "NKT_cells"))
```

### Subset Dataset Again

Second Subset -\> subsetting to remove treatment conditions involving theraputics as this is outside the bounds of the scope of the investigation.

Including:

-   DMG Mice (represented as untreated or UT)

-   Control: No Cancer (Treatment and Disease Naive Group)

-   Surgery Control (SHAM - Injected with Saline - to make sure that the stress of surgery isn't a factor and can be compared between Naive, UT and Sham)

Excluding:

-   ONC201 - Dordaviprone -\> DRD2 antagonist and ClpP agonist

    -   <https://doi.org/10.1158/0008-5472.CAN-23-0186>

    -   <https://doi.org/10.1158/2159-8290.CD-23-0131>

-   Dex - Dexmethasone -\> Corticosteroid

-   Dex & ONC201 Combo

```{r}
seu_NKT_focused <- subset(seu_NKT, subset = treatment %in% c("UT", "NAIVE", "SHAM"))
```
